The Mysterious Lives Of Speckles. I. Residual atmospheric speckle lifetimes behind AO-fed coronagraphs

### 2021-03-21 specklelives.tex
rewrites

\begin{abstract}
High contrast imaging observations are fundamentally limited by the spatially and temporally correlated noise source called speckles.  Suppression and removal of speckles are the key goals of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.   The formation of speckles as spatial components of the instantaneous point-spread function (PSF) is well understood. However, while many empirical and simulation studies have been carried out, there has been relatively little effort focused on a theoretical understanding of the time-domain behavior of the intensity behind a coronagraphic wavefront control system.  Here we present a method for calculating the temporal power spectral density (PSD) of the intensity, and then show how the PSD can be used to calculate the statistical speckle lifetime.  Considering a frozen-flow turbulence model, we use these techniques to analyze the speckle lifetimes behind a MagAO-X-like AO system and extremely large telescope (ELT) scale systems.   We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on MagAO-X.  Speckle-limited integration time is directly proportional to these lifetimes.  While the lifetime varies with telescope diameter, windspeed, and turbulence strength, there are no simple scaling laws among these quantities.   Finally, we describe how the intensity PSD can be used to explore the fundamental limits of post-processing in high contrast imaging.  These results motivate continued effort in the area of wavefront sensor and telemetry based reconstruction and speckle sensing for PSF subtraction, which promise to eliminate the long temporal correlations which currently limit high contrast imaging.
\end{abstract}

\section{Introduction}
Characterization of extrasolar planets with resolved imaging is limited by the halo of starlight scattered onto the image plane by optical aberrations.  In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.  This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in \ogrmv{a}\ogadd{long exposure} coronagraphic observation\ogadd{s}.    Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.   For \ogadd{ground-based} telescopes \ogrmv{on the Earth's surface}, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence \ogrmv{in the Earth's atmosphere}.

@@

In reality, the temporal behavior of atmospheric speckles falls between these regimes. These speckles vary on short time-scales \ogrmv{making standard PSF-subtraction essentially impossible}\ogadd{so they do are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.  The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.

There have been several definitions of speckle lifetime presented in the literature.  Early treatments of speckle dynamics were motivated by speckle interferometry, which relies on exposures short enough to ``freeze the atmosphere''.  The optimum exposure time is therefore dependent on the coherence time of the intensity in the focal plane.  A common definition for this coherence time was based on a $1/e$ reduction in the autocorrelation of the intensity, which was experimentally found to be on the order of a few ms \citep{1978ApOpt..17.3779S,1990JMOp...37.1247D}.    \citet{1982JOpt...13..263R} derived the following expression for the autocorrelation e-folding time in multi-layer Kolmogorov atmosphere:
\begin{equation}
\tau_{boil} = 0.36 \frac{r_0}{\sqrt{\sigma^2_\mathcal{V}}}
\label{eqn:e-fold}
\end{equation}
called there the ``speckle boiling time'', where $r_0$ is Fried's parameter and $\sigma^2_\mathcal{V}$ is the variance with respect to height of the layer wind velocities $\mathcal{V}$  above the telescope.  At good astronomical sites \ogrmv{such as MKO and LCO}, values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms. \citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and showed that speckles dominate photon noise by factors of 100 or more in high contrast imaging.

The autocorrelation e-folding time is a somewhat arbitrary choice.  If we instead consider the evolution of the variance in the estimate of the mean given uncorrelated trials, we can define the speckle lifetime according to
\begin{equation}
\sigma_{mean}^2 = \frac{\sigma_o^2}{t/\tau}
\label{eqn:varmean_def_intro}
\end{equation}
Here $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.  Note that this assumes $t >> \tau$.  \citet{1986JOSAA...3.1001A} employed this definition, and argued for longer speckle lifetimes, replacing $0.36$ by $~$$1.14$ in Equation (\ref{eqn:e-fold}) based on arguments similar to \citet{1982JOpt...13..263R}.  Using AO corrected images, \citet{2006ApJ...637..541F} empirically found significantly longer values of $\tau$, giving an overall median of 175 ms at Lick observatory.  Of note, the results of \citet{2006ApJ...637..541F} hinted that inside the region of the image plane controlled by the AO system, the speckle lifetime was significantly shorter.

\citet{2005SPIE.5903..170M} considered the time averaging of residual atmospheric speckles behind a coronagraph, directly analyzing the reduction in variance as a function of exposure time.  They argued that the speckle lifetime should not be affected by the properties of the AO system nor by the seeing parameter $r_0$, and instead should be determined by the wind-crossing time of a telescope with diameter $D$.  Based on a simulation, \citet{2005SPIE.5903..170M} found that 
\begin{equation}
\tau \approx 0.3 \frac{D}{\mathcal{V}}
\end{equation}
for use in Equation (\ref{eqn:varmean_def_intro}).  This result is quite different from the previous predictions, but coarsely in agreement with the experiments by \citet{2006ApJ...637..541F} (it is challenging to make more definitive experimental comparisons due to the uncertainty in atmospheric parameters).

@@

In the present study, we develop a semi-analytic method for predicting the residual speckle lifetime according to Equation (\ref{eqn:varmean_def_intro}).  This model explicitly accounts for closed-loop AO control, including advanced predictive control laws.  We verify the results of the model by comparison to end-to-end simulations of an ideal closed-loop AO system.  We here concentrate on the speckles due to the atmosphere under the frozen flow hypothesis, and do not consider more complicated turbulence evolution (e.g. boiling) nor do we consider instrumental speckles. We begin with a review of the theory of speckle noise in coronagraph images, followed by a derivation of the statistical speckle lifetime based on the power spectral density (PSD) of the intensity.  We then extend the methods of \citet{2018JATIS...4a9001M} to calculate the intensity PSD.  The model predictions for contrast and speckle lifetimes are compared, and we use the model to predict the behavior of specle lifetimes under various conditions.  Finally, we use the intensity PSDs to consider the benefits of telemetry-based post-processing.

\section{Noise Dynamics in Coronagraphic Focal Planes}

The quantity of interest in this analysis is the intensity in the focal plane as a function of time, $I(\vec{r}, t)$, where $\vec{r}$ is the position.  Useful units for this analysis are photons/sec/$(\lambda/D)^2$.  Following \citet{2007ApJ...669..642S} we consider three contributions to the intensity, such that
\begin{equation}
I(\vec{r},t) = I_{*}(t)\left[I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t)\right]
\end{equation}
where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, itself a function of time due to Strehl ratio variation.  $I_c$ is the fractional contribution from residual diffraction from the coronagraph, as well as purely static aberrations.  $I_{as}$ is the fractional contribution from residual atmospheric speckles, and $I_{qs}$ is the fractional contribution from quasi-static (QS) speckles.  The instantaneous raw contrast is
\begin{equation}
C(\vec{r},t) = \frac{I(\vec{r},t)}{I_*(t)} = I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t).
\end{equation}

The total variance, considering both speckles and and photon noise, \jrmadd{at a point in the focal plane (fp)} is \citep{2007ApJ...669..642S} 
\begin{equation}
\sigma^2_{fp}  = I_* \Delta t [\underbrace{I_c + I_{as} + I_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(I_{as}^2 + 2[I_cI_{as} + I_{as}I_{qs}]  \right)}_{\mbox{atm. speckles}} +  \underbrace{\tau_{qs}\left( I_{qs}^2 + 2 I_c I_{qs}\right)}_{\mbox{QS speckles}}]]
\label{eqn:variance}
\end{equation}
where $\tau_{as}$ is the lifetime of the residual atmospheric speckles, and $\tau_{qs}$ is the lifetime of the QS speckles. $\Delta t$ is the exposure time.  For simplicity we have suppressed the $\vec{r}$ dependence.

@@

Our present focus is on the atmospheric residual $I_{as}$.  From here on we will assume a perfect coronagraph and instrument, such that $I_c = 0$ and $I_{qs}$ = 0.  These (optimistic) conditions eliminate the speckle pinning \citep{2001ApJ...558L..71B} terms in Equation (\ref{eqn:variance}).  The variance is then
\begin{equation}
\sigma^2_{fp}  = I_* \Delta t \left(I_{as} + I_* \tau_{as} I_{as}^2 \right)
\label{eqn:variance_as}
\end{equation}
If we attempt to observe a planet, with a contrast with respect to the star of $C_p$, then the peak signal-to-noise ratio of the planet detection will be
\begin{equation}
\mbox{S/N} = \frac{C_p \sqrt{I_* \Delta t}}{ \sqrt{I_{as} (1 + I_*I_{as} \tau_{as})}} 
\end{equation}
For argument's sake, assume a star having $I_* = 10^8$ photon\ogrmv{s}/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $I_{as} = 1\times10^{-4}$ (see \citet{2018JATIS...4a9001M}).  Next we (for now) assume the \citet{2005SPIE.5903..170M} crossing-time speckle lifetime for a 10 m telescope with 10 m/s wind, giving $\tau_{as}$ = 0.3 s.  We now have $I_*I_{as} \tau_{as} = 3000$.  We see that under such conditions  $I_*I_{as} \tau_{as} >> 1$, which would hold even for significantly smaller $\tau_{as}$.  This leads us to
\begin{equation}
\mbox{S/N} \approx \frac{C_p \sqrt{\Delta t}}{ I_{as} \sqrt{\tau_{as}}}. 
\end{equation}
Such an observation is therefore speckle noise limited, and the S/N improves with square-root of the number of realizations of the speckle pattern as $\sqrt{\Delta t/ \tau_{as}}$.  The ratio of the time to reach a given S/N when speckle limited, $\Delta t_{sl}$, to the time needed when photon noise limited, $\Delta t_{pn}$ is
\begin{equation}
\frac{\Delta t_{sl}}{\Delta t_{pn}} = I_*I_{as} \tau_{as} >> 1.
\label{eqn:relexptime}
\end{equation}

@@

\subsection{Speckle Lifetime From A PSD}
\label{sec:psd_lifetime}
We will denote the temporal PSD as $\mathcal{T}(f)$ where $f$ is frequency.  We will restrict ourselves to one-sided PSDs, that is $f \ge 0$.  The PSD is normalized such that the variance of measurements from a process governed by the PSD is
\begin{equation}
\sigma_o^2 = \int_{0}^{f_s/2} \mathcal{T}(f) df.
\label{eqn:process_var}
\end{equation}
where $f_s$ is the sampling frequency defined by the sampling rate $\Delta t$ according to $f_s = 1/\Delta t$.

According to the Wiener-Khinchin theorem, the PSD \jrmadd{is the Fourier transform of the autocorrelation} \jrmrmv{(they are a Fourier transform pair)}.  This means that the PSD describes how the measurements derived from the process are correlated. \jrmadd{\citet{1976JOSA...66..207N} exploited this concept to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov phase screens.  Here we adapt that analysis to arbitrary 1-dimensional PSDs using the Legendre polynomials.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Speckle Intensity PSDs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Speckle Intensity PSDs}

\subsection{Intensity as a Function of Time}
Here we very briefly review the derivation of the intensity time-series from \citet{2018JATIS...4a9001M}, with the goal to introduce the notation needed for this analysis.  See that paper for a detailed treatment.

@@

\subsection{A Recipe for Intensity PSDs}
\label{sect:recipe}
We are now ready to calculate $\mathcal{T}_{I_{mn}}(f)$.  Unfortunately, we have found no purely analytic way to calculate $\langle \left| \mathcal{F}_t\left\{ h^2 \right\} \right|^2 \rangle$ given only $\langle \left| \mathcal{F}_t\left\{ h \right\} \right|^2 \rangle$. \jrmadd{ This is further complicated by the spatial correlations and temporal relationship between modes.}  We therefore employ a straightforward Monte Carlo procedure:
\begin{enumerate}
\item Generate an open-loop time-series for each $h_{mn}^p$ for both parities.  \jrmadd{Note that $h_{mn}^-$ is needed for the spatial correlation step, even though it is discarded eventually.} \label{step:olh}
\begin{enumerate}
\item Generate a Gaussian white-noise time-series of the desired length
\item Fourier transform the white-noise time-series
\item Multiply the Fourier transform by the square-root of the open-loop PSD.  We now have $\mathcal{F}_t\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
\item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$, and then normalize to unit variance.
\item Repeat for each mode.
\end{enumerate}
\item Correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix.  See \citet{2018JATIS...4a9001M} for a treatment of the covariance matrix calculation for the Fourier basis.
\item For each spatial frequency $mn$, replace $h_{mn}^-(t)$ with the phase shifted $h_{mn}^+(t)$ as described above.
\item Generate a WFS measurement noise time-series for each mode, $\eta_{mn}^+$ and $\eta_{mn}^-$, and Fourier transform them.  These are white-noise PSDs, with no phase relationship.
\item Generate the closed-loop Fourier transforms for each mode \label{step:clh}
\begin{eqnarray}
\mathcal{F}_t\{h_{mn,cl}^+\} = \mathcal{F}_t\{h_{mn}^+\} \mathrm{ETF}(f) + \mathcal{F}_t\{\eta_{mn}^+\} \mathrm{NTF}(f) \\
\mathcal{F}_t\{h_{mn,cl}^-\} = \mathcal{F}_t\{h_{mn}^-\} \mathrm{ETF}(f) + \mathcal{F}_t\{\eta_{mn}^-\} \mathrm{NTF}(f) \nonumber
\end{eqnarray}
\item Calculate the closed-loop modal coefficient time-series with the inverse Fourier transform.  We now have $h_{mn,cl}^+(t)$ and $h_{mn,cl}^-(t)$ 
\item Now we can calculate the focal plane intensity as a function of time
\begin{equation}
I_{mn}(t) \approx \frac{2\pi}{\lambda}\left[ \left(h_{mn,cl}^+\right)^2 + \left(h_{mn,cl}^-\right)^2 \right] 
\end{equation}
\item Finally, we calculate $|\mathcal{F}_t\{I_{mn}(t)\}|^2$.  For best results this is done with a window function applied.
\item Now the entire procedure is repeated N times, averaging the result to form the estimate of $\mathcal{T}_{I_{mn}}(f)$.
\end{enumerate}

@@

\section{Results}

@@

\subsection{Simulations}
\label{sect:simulations}
We conducted a series of end-to-end numerical simulations of an idealized MagAO-X-like instrument to validate our calculations.  Our goal is not to develop a realistic hardware simulation, rather to ensure that the the calculations produce reasonable results.

@@

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=3.54in]{contrast0mag.pdf}
\includegraphics[width=3.54in]{contrast8mag.pdf}
\caption{Comparison of long-exposure raw contrast from semi-analytic calculations (Section \ref{sect:recipe}) and end-to-end simulation (Section \ref{sect:simulations}) for a 6.5 m telescope. Left: without photon noise, considering the input turbulence only.  Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS. Note that the colorbar is adjusted to best show detail.  Agreement between the calculations and simulations is very good.\label{fig:contrast_2Dcomp}}
\end{figure}

@@

\subsection{Trends in Speckle Lifetime}
\label{sec:trends}

>>>>

\begin{abstract}
High contrast imaging observations are fundamentally limited by the spatially and temporally correlated noise source called speckles.  Suppression of speckles is the key goal of wavefront control and adaptive optics (AO), coronagraphy, and a host of post-processing techniques.  Here we present a method for calculating the temporal power spectral density (PSD) of the intensity, and then show how the PSD can be used to calculate the statistical speckle lifetime.  Speckle-limited integration time is directly proportional to this lifetime.  Considering a frozen-flow turbulence model, we analyze the speckle lifetimes behind a MagAO-X-like AO system and extremely large telescope (ELT) scale systems.    We find that standard AO control shortens atmospheric speckle lifetime from $\sim$130 ms to $\sim$50 ms, and predictive control will further shorten the lifetime to $\sim$20 ms on MagAO-X.  Finally, we use these results to describe how ground-based telescopes will achieve the photon-noise limit in high contrast imaging.
\end{abstract}

%\jrmrmv{The formation of speckles as spatial components of the instantaneous point-spread function (PSF) is well understood. However, while many empirical and simulation studies have been carried out, there has been relatively little effort focused on a theoretical understanding of the time-domain behavior of the intensity behind a coronagraphic wavefront control system.}

\section{Introduction}
Characterization of extrasolar planets with resolved imaging is limited by the halo of starlight scattered onto the image plane by optical aberrations.  In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.  This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in \ogrmv{a}\ogadd{long exposure} coronagraphic observation\ogadd{s}.    Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.   For \ogadd{ground-based} telescopes \ogrmv{on the Earth's surface}, the aberrations \ogadd{also} include those imposed by \ogadd{atmospheric} turbulence. \ogrmv{in the Earth's atmosphere}

@@

In reality, the temporal behavior of atmospheric speckles falls between these regimes. These speckles vary on short time-scales \ogrmv{making standard PSF-subtraction essentially impossible}\ogadd{so they are not removed by standard PSF subtraction}, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.  The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.

\subsection{Atmospheric Coherence}
There have been several definitions of speckle lifetime presented in the literature.  Early treatments of speckle dynamics were motivated by speckle interferometry, which relies on exposures short enough to ``freeze the atmosphere''.  The optimum exposure time is therefore dependent on the coherence time of the intensity in the focal plane.  A common definition for this coherence time was based on a $1/e$ reduction in the autocorrelation of the intensity, which was experimentally found to be on the order of a few ms \citep{1978ApOpt..17.3779S,1990JMOp...37.1247D}.    \citet{1982JOpt...13..263R} derived the following expression for the autocorrelation e-folding time in a multi-layer Kolmogorov atmosphere:
\begin{equation}
\tau_{boil} = 0.36 \frac{r_0}{\sqrt{\sigma^2_\mathcal{V}}}
\label{eqn:e-fold}
\end{equation}
called there the ``speckle boiling time'', where $r_0$ is Fried's parameter and $\sigma^2_\mathcal{V}$ is the variance with respect to height of the layer wind velocities $\mathcal{V}$  above the telescope.  At good astronomical sites \ogrmv{such as MKO and LCO}, values of $\tau_{boil}$ should range from $\sim$5 to $\sim$20 ms. \citet{1999PASP..111..587R} employed Equation (\ref{eqn:e-fold}) for the speckle lifetime and showed that speckles dominate photon noise by factors of 100 or more in high contrast imaging.

\subsection{Variance of the Mean}
The autocorrelation e-folding time is a somewhat arbitrary choice.  If we instead consider the evolution of the variance in the estimate of the mean given uncorrelated trials, we can define the speckle lifetime according to
\begin{equation}
\sigma_{mean}^2 = \frac{\sigma_o^2}{t/\tau}
\label{eqn:varmean_def_intro}
\end{equation}
\jrmadd{This simply states that the error in a measurement of the mean improves with the number of uncorrelated trials $t/\tau$.  Defined this way,} $\tau$ is given by the integral of the autocorrelation \citep{2006ApJ...637..541F}, and in this expression determines the number of uncorrelated realizations of the speckle pattern included in the mean.  Note that this assumes $t >> \tau$.  \citet{1986JOSAA...3.1001A} employed this definition, and argued for longer speckle lifetimes, replacing $0.36$ by $~$$1.14$ in Equation (\ref{eqn:e-fold}) based on arguments similar to \citet{1982JOpt...13..263R}.  Using AO corrected images, \citet{2006ApJ...637..541F} empirically found significantly longer values of $\tau$, giving an overall median of 175 ms at Lick observatory.  Of note, the results of \citet{2006ApJ...637..541F} hinted that inside the region of the image plane controlled by the AO system, the speckle lifetime was significantly shorter.

\citet{2005SPIE.5903..170M} considered the time averaging of residual atmospheric speckles behind a coronagraph, directly analyzing the reduction in variance as a function of exposure time.  They argued that the speckle lifetime should not be affected by the properties of the AO system nor by the seeing parameter $r_0$, and instead should be determined by the wind-crossing time of a telescope with diameter $D$.  Based on a simulation, \citet{2005SPIE.5903..170M} found that 
\begin{equation}
\tau \approx 0.3 \frac{D}{\mathcal{V}}
\end{equation}
for use in Equation (\ref{eqn:varmean_def_intro})\footnote{\citet{2005SPIE.5903..170M} reported a speckle decorrelation time which would be a factor of two longer than the statistical lifetime}.  This result is quite different from the previous predictions, but coarsely in agreement with the experiments by \citet{2006ApJ...637..541F} (it is challenging to make more definitive experimental comparisons due to the uncertainty in atmospheric parameters).

@@

\subsection{Goals of This Study}
In this study, we seek to describe the time-domain behavior of speckle noise in AO-fed coronagraphic images and its impact on sensitivity from as close to first principles as possible.  This requires a traversal of intensity statistics, signal processing, wavefront propagation, and control theory, and so is somewhat involved. To accomplish our goals we will work through the following tasks, organized by section:
\begin{itemize}
\item Section \ref{sec:noise}: Review speckle statistics to understand how atmospheric speckles dramatically reduce sensitivity. \\
\item Section \ref{sec:psd_lifetime}: Develop an analytic framework for analyzing power spectral densities (PSDs), which allows us to derive an expression for the statistical (mean-averaging) lifetime from a PSD.
\item Section \ref{sec:intensity}: Show how to calculate the intensity PSD in the coronagraphic focal plane behind a closed-loop AO system with a semi-analytic model.
\item Section \ref{sec:results}: Employ an end-to-end AO system simulation to (a) validate the statistical lifetime calculations and (b) assess the performance of the semi-analytic model.
\item Section \ref{sec:trends}: Use the analytic framework to analyze speckle lifetime over a range of telescope diameters and atmospheric parameters.
\item Section \ref{sec:reconstruction}: We propose a method to analyze post-processing in the time-domain and discuss its implications for the limits of ground-based high-contrast imaging sensitivity.
\item Section \ref{sec:discussion}: We then discuss various implications of these results.
\end{itemize}
For the impatient reader, Sections \ref{sec:psd_lifetime} through \ref{sec:results} and Section \ref{sec:reconstruction} contain the technical details.  The key points can be found in Sections \ref{sec:noise}, \ref{sec:trends}, and \ref{sec:discussion}.

%In the present study, we develop a semi-analytic method for predicting the residual speckle lifetime according to Equation (\ref{eqn:varmean_def_intro}).  This model explicitly accounts for closed-loop AO control, including advanced predictive control laws.  We verify the results of the model by comparison to end-to-end simulations of an ideal closed-loop AO system.  We here concentrate on the speckles due to the atmosphere under the frozen flow hypothesis, and do not consider more complicated turbulence evolution (e.g. boiling) nor do we consider instrumental speckles. We begin with a review of the theory of speckle noise in coronagraph images, followed by a derivation of the statistical speckle lifetime based on the power spectral density (PSD) of the intensity.  We then extend the methods of \citet{2018JATIS...4a9001M} to calculate the intensity PSD.  The model predictions for contrast and speckle lifetimes are compared, and we use the model to predict the behavior of specle lifetimes under various conditions.  Finally, we use the intensity PSDs to consider the benefits of telemetry-based post-processing.

\section{Noise Dynamics in Coronagraphic Focal Planes}
\label{sec:noise}
\jrmadd{The key metric for sensitivity in a high-contrast imaging observation is the signal-to-noise ratio, S/N.  The signal we collect from a planet, which has planet to star contrast of $C_p$, is }
\begin{equation}
\mbox{S}_p = I_{*}(t) C_p \Delta t
\end{equation}
\jrmadd{where $I_{*}$ is the peak intensity of the star's PSF without a coronagraph, which is a function of time due to Strehl ratio variation, and $\Delta t$ is the exposure time.  Here we will assume units of photons/$(\lambda/D)^2$/second for the intensity.}

\jrmadd{The dominant noise source in such an observation is the halo of scattered starlight underlying the planet signal.  At the planet position $\vec{r}_p$ this has intensity $I_h(\vec{r}_p,t)$. Following \citet{2007ApJ...669..642S} we consider three contributions to the halo intensity:}
\begin{equation}
I_h(\vec{r}_p,t) = I_{*}(t)\left[I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t)\right]
\end{equation}
$I_c$ is the fractional contribution from residual diffraction from the coronagraph, as well as purely static aberrations.  $I_{as}$ is the fractional contribution from residual atmospheric speckles, and $I_{qs}$ is the fractional contribution from quasi-static (QS) speckles.  The halo contributes noise in two distinct ways: photon noise and correlated speckle noise.
%From this we can state the instantaneous raw halo contrast
%\begin{equation}
%C_h(\vec{r},t) = \frac{I(\vec{r},t)}{I_*(t)} = I_c(\vec{r},t) + I_{as}(\vec{r},t) + I_{qs}(\vec{r},t).
%\end{equation}
The total variance, considering both speckles and and photon noise, \jrmadd{under the planet signal} is \citep{2007ApJ...669..642S} 
\begin{equation}
\sigma^2_{p}  = I_* \Delta t [\underbrace{I_c + I_{as} + I_{qs}}_{\mbox{photon noise}} + I_* [ \underbrace{\tau_{as}\left(I_{as}^2 + 2[I_cI_{as} + I_{as}I_{qs}]  \right)}_{\mbox{atm. speckles}} +  \underbrace{\tau_{qs}\left( I_{qs}^2 + 2 I_c I_{qs}\right)}_{\mbox{QS speckles}}]]
\label{eqn:variance}
\end{equation}
where $\tau_{as}$ is the lifetime of the residual atmospheric speckles, and $\tau_{qs}$ is the lifetime of the QS speckles.  For simplicity we have suppressed the $\vec{r}$ dependence.

@@

Our present focus is on the atmospheric residual $I_{as}$.  From here on we will assume a perfect coronagraph and instrument, such that $I_c = 0$ and $I_{qs}$ = 0.  These (optimistic) conditions eliminate the speckle pinning \citep{2001ApJ...558L..71B} terms in Equation (\ref{eqn:variance}).  The variance is then
\begin{equation}
\sigma^2_{p}  = I_* \Delta t \left(I_{as} + I_* \tau_{as} I_{as}^2 \right)
\label{eqn:variance_as}
\end{equation}

Now the peak signal-to-noise ratio of the planet detection will be
\begin{equation}
\mbox{S/N} = \frac{C_p \sqrt{I_* \Delta t}}{ \sqrt{I_{as} (1 + I_*I_{as} \tau_{as})}} 
\end{equation}
For argument's sake, assume a star having $I_* = 10^8$ photon\ogrmv{s}/sec/$(\lambda/D)^2$ in the peak (roughly an 8th mag star on a 10 m telescope in the near-IR), and $I_{as} = 1\times10^{-4}$ (see \citet{2018JATIS...4a9001M}).  Next we (for now) assume the \citet{2005SPIE.5903..170M} crossing-time speckle lifetime for a 10 m telescope with 10 m/s wind, giving $\tau_{as}$ = 0.3 s.  We now have $I_*I_{as} \tau_{as} = 3000$.  We see that under such conditions  $I_*I_{as} \tau_{as} >> 1$, which would hold even for significantly smaller $\tau_{as}$.  This leads us to
\begin{equation}
\mbox{S/N} \approx \frac{C_p \sqrt{\Delta t}}{ I_{as} \sqrt{\tau_{as}}}. 
\end{equation}
Such an observation is therefore speckle noise limited, and the S/N improves with square-root of the number of realizations of the speckle pattern as $\sqrt{\Delta t/ \tau_{as}}$.  The ratio of the time to reach a given S/N when speckle limited, $\Delta t_{sl}$, to the time needed when photon noise limited, $\Delta t_{pn}$ is
\begin{equation}
\frac{\Delta t_{sl}}{\Delta t_{pn}} = I_*I_{as} \tau_{as} >> 1.
\label{eqn:relexptime}
\end{equation}

@@

\section{Speckle Lifetime From A PSD}
\label{sec:psd_lifetime}
We will denote the temporal PSD as $\mathcal{T}(f)$ where $f$ is frequency.  We will restrict ourselves to one-sided PSDs, that is $f \ge 0$.  The PSD is normalized such that the variance of measurements from a process governed by the PSD is
\begin{equation}
\sigma_o^2 = \int_{0}^{f_s/2} \mathcal{T}(f) df.
\label{eqn:process_var}
\end{equation}
where $f_s$ is the sampling frequency defined by the sampling rate $\Delta t$ according to $f_s = 1/\Delta t$.

According to the Wiener-Khinchin theorem, the PSD \jrmadd{is the Fourier transform of the autocorrelation} \jrmrmv{(they are a Fourier transform pair)}.  This means that the PSD describes how the measurements derived from the process are correlated. \jrmadd{\citet{1976JOSA...66..207N} exploited this concept in two dimensions to derive the coefficient covariances in the Zernike polynomial expansion of Kolmogorov phase screens.  Here we adapt that analysis to arbitrary one dimensional PSDs using the Legendre polynomials.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Speckle Intensity PSDs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Speckle Intensity PSDs}
\label{sec:intensity}
\subsection{Intensity as a Function of Time}
Here we very briefly review the derivation of the intensity time-series from \citet{2018JATIS...4a9001M}, with the goal to introduce the notation needed for this analysis.  See that paper for a detailed treatment.

@@

\subsection{A Recipe for Intensity PSDs}
\label{sec:recipe}
We are now ready to calculate $\mathcal{T}_{I_{mn}}(f)$.  Unfortunately, we have found no purely analytic way to calculate $\langle \left| \mathcal{F}_t\left\{ h^2 \right\} \right|^2 \rangle$ given only $\langle \left| \mathcal{F}_t\left\{ h \right\} \right|^2 \rangle$. \jrmadd{ This is further complicated by the spatial correlations and temporal relationship between modes.}  We therefore employ a straightforward Monte Carlo procedure:
\begin{enumerate}
\item Generate an open-loop time-series for each $h_{mn}^p$ for both parities.  \jrmadd{Note that $h_{mn}^-$ is needed for the spatial correlation step, even though it is discarded eventually.} \label{step:olh}
\begin{enumerate}
\item Generate a Gaussian white-noise time-series of the desired length
\item Fourier transform the white-noise time-series
\item Multiply the Fourier transform by the square-root of the open-loop PSD.  We now have $\mathcal{F}_t\left\{h_{mn}^p\right\}$ \citep[cf.][]{kasdin_falpha}.
\item Apply the inverse Fourier transform to produce $h_{mn}^p(t)$, and then normalize to unit variance.
\item Repeat for each mode.
\end{enumerate}
\item Correlate the modal time-series by multiplying the vector of coefficients at each time-step by the decomposition of the covariance matrix.  See \citet{2018JATIS...4a9001M} for a treatment of the covariance matrix calculation for the Fourier basis.
\item For each spatial frequency $mn$, replace $h_{mn}^-(t)$ with the phase shifted $h_{mn}^+(t)$ as described above.
\item Generate a WFS measurement noise time-series for each mode, $\eta_{mn}^+$ and $\eta_{mn}^-$, and Fourier transform them.  These are white-noise PSDs, with no phase relationship.
\item Generate the closed-loop Fourier transforms for each mode \label{step:clh}
\begin{eqnarray}
\mathcal{F}_t\{h_{mn,cl}^+\} = \mathcal{F}_t\{h_{mn}^+\} \mathrm{ETF}(f) + \mathcal{F}_t\{\eta_{mn}^+\} \mathrm{NTF}(f) \\
\mathcal{F}_t\{h_{mn,cl}^-\} = \mathcal{F}_t\{h_{mn}^-\} \mathrm{ETF}(f) + \mathcal{F}_t\{\eta_{mn}^-\} \mathrm{NTF}(f) \nonumber
\end{eqnarray}
\item Calculate the closed-loop modal coefficient time-series with the inverse Fourier transform.  We now have $h_{mn,cl}^+(t)$ and $h_{mn,cl}^-(t)$ 
\item Now we can calculate the focal plane intensity as a function of time
\begin{equation}
I_{mn}(t) \approx \frac{2\pi}{\lambda}\left[ \left(h_{mn,cl}^+\right)^2 + \left(h_{mn,cl}^-\right)^2 \right] 
\end{equation}
\item Finally, we calculate $|\mathcal{F}_t\{I_{mn}(t)\}|^2$.  For best results this is done with a window function applied.
\item Now the entire procedure is repeated N times, averaging the result to form the estimate of $\mathcal{T}_{I_{mn}}(f)$.
\end{enumerate}

@@

\section{Results}
\label{sec:results}

@@

\subsection{Simulations}
\label{sec:simulations}
We conducted a series of end-to-end numerical simulations of an idealized MagAO-X-like instrument to validate our calculations.  Our goal is not to develop a realistic hardware simulation, rather to ensure that the the calculations produce reasonable results.

@@

\begin{figure}
\hspace{-0.3in}
\includegraphics[width=3.54in]{contrast0mag.pdf}
\includegraphics[width=3.54in]{contrast8mag.pdf}
\caption{Comparison of long-exposure raw contrast from semi-analytic calculations (Section \ref{sec:recipe}) and end-to-end simulation (Section \ref{sec:simulations}) for a 6.5 m telescope. Left: without photon noise, considering the input turbulence only.  Right: with photon noise on an 8th mag star assuming the sensitivity of an unmodulated pyramid WFS. Note that the colorbar is adjusted to best show detail.  Agreement between the calculations and simulations is very good.\label{fig:contrast_2Dcomp}}
\end{figure}

@@

\section{Trends in Speckle Lifetime}
\label{sec:trends}
